Embeddings and Attention in Predictive Modeling

### 2022-03-29 manuscript/manuscript.Rmd
color changes in blue

If $K$ is set equal to $1$, then Equation \ref{NN} reduces to nothing more than a GLM. A neural network with $K=2$ is called a shallow neural network and for $K \geq 2$, the network is called a deep neural network. The matrix $X$ of data input to the network can be composed of both continuous variables as well as categorical variables, which can be pre-processed using one-hot or dummy encoding. As mentioned above, a different option is to use encodings, which we discuss in more detail next.

@@

Since we only have a few parameters in this model, we exhibit the full list of fitted coefficients in Table \@ref(tab:glm2).

@@

| Model                                               |   RMSE |    MAE |
|:----------------------------------------------------|-------:|-------:|
| Model 4: MLP with multidimensional embeddings       | 60,973 | 37,574 |
| Model 5: Simple Attention                           | 60,601 | 36,739 |
| Model 6: TabTransformer                             | 59,900 | 36,343 |
| Model 7: TabNet                                     | 62,938 | 38,386 |
| Linear regression, predictions capped below at 0.01 | 60,879 | 38,431 |

>>>>

If $K$ is set equal to $1$, then Equation \ref{NN} reduces to a \textcolor{blue}{form similar to the GLM, where $\sigma$ is analogous to the link function and $a_1X+b_1$ the linear predictor.} A neural network with $K=2$ is called a shallow neural network and for $K \geq 2$, the network is called a deep neural network. The matrix $X$ of data input to the network can be composed of both continuous variables as well as categorical variables, which can be pre-processed using one-hot or dummy encoding. As mentioned above, a different option is to use encodings, which we discuss in more detail next.

@@

\textcolor{blue}{Regardless of the hyperparameters chosen, care needs to be taken in ensuring the integrity of the model validation process, since when we train the embeddings, we utilize the information from the response variables. Therefore, the holdout data must not be seen by the neural network training process or the GLM fitting process.}

\textcolor{blue}{Since we only have a few parameters in this model, we exhibit the full list of fitted coefficients in Table} \@ref(tab:glm2). 
\textcolor{blue}{For each categorical variable, there is only one coefficient estimate. This is because each categorical variable has been replaced by a
single scalar value, so we treat it as if it were any other continuous predictor.}

\textcolor{blue}{One interesting ramification of our choice of embedding dimension of 1 in this example is that we can compute the "implied relativities" of the factor levels in the new GLM. Since each level corresponds to a unique scalar value, we can obtain its contribution to the linear predictor by multiplying by the coefficient estimate of the variable. For example, multiplying the estimate of -0.075 for} `occupancy_type` in Table \@ref(tab:glm2) \textcolor{blue}{by each value in Table} \@ref(tab:embeddings-flood-zone), \textcolor{blue}{we obtain the unnormalized contributions of each level, which can then be exponentiated and normalized with respect to the base level} `Single family residence` \textcolor{blue}{in order to obtain the implied relativities, which could then be compared with the first GLM's output in Table} \@ref(tab:relativities).

@@

| Model                                               |   RMSE |    MAE |
|:----------------------------------------------------|-------:|-------:|
| Model 4: MLP with multidimensional embeddings       | 60,973 | 37,574 |
| Model 5: Simple Attention                           | 60,601 | 36,739 |
| \textcolor{blue}{Model 6: TabTransformer}                             | 59,900 | 36,343 |
| \textcolor{blue}{Model 7: TabNet}                                     | 62,938 | 38,386 |
| Linear regression, predictions capped below at 0.01 | 60,879 | 38,431 |
